Interestingness Measures for Theory Exploration

### 2015-11-02 TransferReport/report.tex
Updates

@@

\subsubsection{Relevance}

@@

\subsubsection{Clustering}
% TODO: ML4PG
% TODO: ACL2(ml)

@@

% TODO:
\cite{Montano-Rivas.McCasland.Dixon.ea:2012}
\cite{Piantadosi.Tenenbaum.Goodman:2012}
\cite{Wille:2005}
\cite{colton1999automatic}
\cite{colton2000agent}
\cite{colton2012automated}
\cite{lenat1977automated}
\cite{mullerunderstanding}
\cite{Bundy.Cavallo.Dixon.ea:2015}
\cite{johansson2009isacosy}
\cite{spector2008genetic}
\cite{colton2012automated}
% TODO: How does https en.wikipedia.org/wiki/Discovery system relate?

@@

  \item{Counterexample Generation}: As mentioned in \S \ref{tests}, there are many Haskell property checkers, which differ mostly in the way they generate values for testing. Since our theory exploration tools are built on these property checkers, we rely on the existence of value generators for all of the types in our theory; yet existing approaches to deriving such generators (e.g. as found in the \texttt{derive} package) rely on choose between constructors, recursively. As demonstrated in \S \ref{haskell}, a type's constructors may not be exported, in which case these methods fail; not to mention the exponential complexity of such recursive branching. It seems that a database of inspectable Haskell functions may be useful for solving this problem, although we have not yet performed experiments in this area.
\end{itemize}

@@

We identify the following potential improvements for existing TE systems such as \textsc{HipSpec}:

\begin{description}
\item{Term generation}:
  Enumerating all type-correct terms is a brute-force solution to this question.
  Scalable alternatives to brute-force algorithms are a well-studied area of
  Artificial Intelligence and Machine Learning. In particular, heuristic
  search algorithms like those surveyed in \cite{blum2011hybrid} could be used.
  We could also use Machine Learning methods to identify some sub-set of a given
  theory, to prioritise over the rest.
\item{Interestingness}
  Various alternative interestingness criteria have been proposed, for
  example those surveyed in \cite{geng2006interestingness}. Augmenting or
  replacing the criteria may be useful, for example to distinguish useful
  relationships from incidental coincidences; or to prevent surprising,
  insightful equations from being discarded because they can be simplified.
\end{description}

We are implementing a system called \textsc{ML4HS} to investigate these ideas.
Its current form is a pre-processor for \textsc{QuickSpec} for prioritising
theory elements. Inspired by the use of premise selection
\cite{kuhlwein2012overview} to reduce the search space in ATP,
we select sub-sets of the given theory to explore, chosen to try and keep
together those expressions which combine in interesting ways, and to separate
those which combine in uninteresting ways.

We hypothesise that similarity-based clustering of expressions, inspired by that
of \textsc{ML4PG} \cite{journals/corr/abs-1212-3618} and related work in ACL2
\cite{heras2013proof}, is an effective method for performing this separation.
Future experiments will test this by comparing the throughput of
\textsc{QuickSpec} with and without the \textsc{ML4HS} pre-processor.

>>>>

We identify the following areas for researching potential improvements in existing TE systems such as \textsc{HipSpec}:

\begin{description}
\item{Term generation}:
  Enumerating all type-correct terms is a brute-force solution to this question. Scalable alternatives to brute-force algorithms are a well-studied area of Artificial Intelligence and Machine Learning \iffalse TODO: Related work reference \fi. In particular, heuristic search algorithms like those surveyed in \cite{blum2011hybrid} could be used.
  We could also use Machine Learning methods like those used for \emph{relevance filtering} \ref{relevance} to identify some sub-set of a given theory, to prioritise over the rest.

\item{Interestingness}
  Various alternative interestingness criteria have been proposed, which we survey in \S \ref{relatedwork}. Augmenting or replacing the criteria may be useful, for example to distinguish useful relationships from incidental coincidences; or to prevent surprising, insightful equations from being discarded because they can be simplified.
\end{description}

@@

\subsubsection{Relevance Filtering}
\label{relevance}

% TODO
\cite{kuhlwein2012overview}

@@

\subsubsection{Clustering}
\label{clustering}

% TODO: ML4PG
\cite{journals/corr/abs-1212-3618}
% TODO: ACL2(ml) (also examples section)
\cite{heras2013proof}

@@

% TODO:
\cite{Montano-Rivas.McCasland.Dixon.ea:2012}
\cite{Piantadosi.Tenenbaum.Goodman:2012}
\cite{Wille:2005}
\cite{colton1999automatic}
\cite{colton2000agent}
\cite{colton2012automated}
\cite{lenat1977automated}
\cite{mullerunderstanding}
\cite{Bundy.Cavallo.Dixon.ea:2015}
\cite{johansson2009isacosy}
\cite{spector2008genetic}
\cite{colton2012automated}
 \cite{geng2006interestingness}
% TODO: How does https en.wikipedia.org/wiki/Discovery system relate?

@@

  \item{Counterexample Generation}: As mentioned in \S \ref{tests}, there are many Haskell property checkers, which differ mostly in the way they generate values for testing. Since our theory exploration tools are built on these property checkers, we rely on the existence of value generators for at least some of the types in our theory; yet existing approaches to deriving such generators (e.g. as found in the \texttt{derive} package) rely on choosing an arbitrary constructor then recursing. As demonstrated in \S \ref{haskell}, a type's constructors might not be exported, in which case these methods fail. Also, the exponential complexity of such naive recursion makes these generators unusable in many cases. It seems that a database of inspectable Haskell functions may be useful for solving this problem in a more practical way, although we have not yet performed experiments in this area.
\end{itemize}

@@

\textsc{ML4HS} is our top-level theory exploration framework, combining \textsc{AstPlugin}, \textsc{TreeFeatures}, the \textsc{Weka} machine learning library and \textsc{MLSpec}.

Inspired by the use of premise selection (see \S \ref{relevance}) to reduce the search space in ATP,
we select sub-sets of the given theory to explore, chosen to try and keep together those expressions which combine in interesting ways, and to separate those which combine in uninteresting ways.

We hypothesise that similarity-based clustering of expressions, inspired by those of \textsc{ML4PG}  and related work in \textsc{ACL2(ml)} (see \S \ref{clustering}), is an effective method for performing this separation.

% TODO: Future experiments will test this by comparing the throughput of \textsc{QuickSpec} with and without the \textsc{ML4HS} pre-processor.
